New cardinality estimation algorithms for HyperLogLog sketches

### 2016-09-18 paper/paper.tex
consistent joint hyperloglog estimation formulas in paper and implementation
added rmse ration of inclusion-exclusion principle and maximum likelihood method for joint cardinality estimation

\documentclass[a4paper]{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{svg}
\usepackage{algpseudocode,algorithm,algorithmicx}
\usepackage{float}
\usepackage{url}
\usepackage{booktabs}
\usepackage{array}
\usepackage{paralist}
\usepackage{verbatim}
\usepackage{subfig}
\usepackage{hyperref}
\usepackage{siunitx}
\usepackage{csvsimple}
\usepackage[capitalise,nameinlink,noabbrev]{cleveref}

% SYMBOLS
\newcommand*{\symDefine}[2]{\newcommand{{#1}}{{#2}}}
\symDefine{\symPrecision}{p}
\symDefine{\symRegRange}{q}
\symDefine{\symRegVal}{k}
\symDefine{\symRegValVariate}{K}
\symDefine{\symNumReg}{m}
\symDefine{\symDataItem}{D}
\symDefine{\symBitRepA}{a}
\symDefine{\symBitRepB}{b}
\symDefine{\symIndexI}{i}
\symDefine{\symIndexK}{k}
\symDefine{\symIndexL}{l}
\symDefine{\symIndexJ}{j}
\symDefine{\symCount}{c}
\symDefine{\symCountVariate}{C}
\symDefine{\symAlpha}{\alpha}
\symDefine{\symBeta}{\beta}
\symDefine{\symA}{a}
\symDefine{\symB}{b}
\symDefine{\symKappa}{\kappa}
\symDefine{\symX}{x}
\symDefine{\symPhi}{\varphi}
\symDefine{\symXEstimate}{\hat{\symX}}
\symDefine{\symXNormalized}{\symX'}
\symDefine{\symN}{n}
\symDefine{\symS}{s}
\symDefine{\symY}{y}
\symDefine{\symZ}{z}
\symDefine{\symInteger}{z}
\symDefine{\symIntegerVariate}{Z}
\symDefine{\symCardinality}{n}
\symDefine{\symCardinalityEstimate}{\hat{\symCardinality}}
\symDefine{\symCardinalityRawEstimate}{\symCardinalityEstimate_\text{raw}}
\symDefine{\symError}{\varepsilon}
\symDefine{\symStopDelta}{\delta}
\symDefine{\symStopEpsilon}{\varepsilon}
\symDefine{\symBigO}{\mathcal{O}}
\symDefine{\symExpectation}{\mathbb{E}}
\symDefine{\symProbability}{P}
\symDefine{\symProbabilityMass}{\rho}
\symDefine{\symRegProbability}{\gamma}
\symDefine{\symLikelihood}{\mathcal{L}}
\symDefine{\symPoissonRate}{\lambda}
\symDefine{\symPoissonRateEstimate}{\hat{\symPoissonRate}}
\symDefine{\symFunc}{f}
\symDefine{\symFuncPrime}{g}
\symDefine{\symHelper}{h}
\symDefine{\symHelperApprox}{\tilde{\symHelper}}
\symDefine{\symHelperW}{w}
\symDefine{\symHelperV}{v}
\symDefine{\symHelperU}{u}
\symDefine{\symSetA}{A}
\symDefine{\symSetB}{B}
\symDefine{\symSetS}{S}
\symDefine{\symSetX}{X}
\symDefine{\symSetASuffix}{a}
\symDefine{\symSetBSuffix}{b}
\symDefine{\symSetXSuffix}{x}
\symDefine{\symCountMatrix}{\mathbf{c}}
\symDefine{\symPowerSeriesFunc}{\xi}
\symDefine{\symSmallCorrectionFunc}{\sigma}
\symDefine{\symLargeCorrectionFunc}{\tau}
\symDefine{\symEpsPowerSeriesFunc}{\varepsilon_\symPowerSeriesFunc}

\sisetup{
fixed-exponent = 2,
scientific-notation = fixed,
}

@@

The transformed log-likelihood function can be written as
\begin{multline}
\symFunc(\symPhi_\symSetASuffix,\symPhi_\symSetBSuffix,\symPhi_\symSetXSuffix):=
\log \symLikelihood(
\symNumReg e^{\symPhi_\symSetASuffix},
\symNumReg e^{\symPhi_\symSetBSuffix},
\symNumReg e^{\symPhi_\symSetXSuffix}
\vert
\boldsymbol{\symRegValVariate}_1,
\boldsymbol{\symRegValVariate}_2
)=
\\
\begin{aligned}
&-
\sum_{\symRegVal=0}^{\symRegRange}
\left(
  \symCountVariate^{<}_{1\symRegVal}+
  \symCountVariate^{=}_{\symRegVal}+
  \symCountVariate^{>}_{1\symRegVal}
\right)
\frac{e^{\symPhi_\symSetASuffix}}{2^{\symRegVal}}
+
\left(
  \symCountVariate^{<}_{2\symRegVal}+
  \symCountVariate^{=}_{\symRegVal}+
  \symCountVariate^{>}_{2\symRegVal}
\right)
\frac{e^{\symPhi_\symSetBSuffix}}{2^{\symRegVal}}
+
\left(
  \symCountVariate^{<}_{1\symRegVal}+
  \symCountVariate^{=}_{\symRegVal}+
  \symCountVariate^{<}_{2\symRegVal}
\right)
\frac{e^{\symPhi_\symSetXSuffix}}{2^{\symRegVal}}
\\
&+
\sum_{\symRegVal=1}^{\symRegRange}
\symCountVariate^{<}_{1\symRegVal}
\log\!\left(1-e^{-\frac{e^{\symPhi_\symSetASuffix}
+e^{\symPhi_\symSetXSuffix}
}{2^{\symRegVal}}}\right)
+
\symCountVariate^{<}_{2\symRegVal}
\log\!\left(1-e^{-\frac{e^{\symPhi_\symSetBSuffix}
+
e^{\symPhi_\symSetXSuffix}
}{2^{\symRegVal}}}\right)
\\
&+
\sum_{\symRegVal=1}^{\symRegRange+1}
\symCountVariate^{>}_{1\symRegVal}
\log\!\left(1-e^{-\frac{e^{\symPhi_\symSetASuffix}
}{2^{\min\left(\symRegVal,\symRegRange\right)}}}\right)
+
\symCountVariate^{>}_{2\symRegVal}
\log\!\left(1-e^{-\frac{e^{\symPhi_\symSetBSuffix}
}{2^{\min\left(\symRegVal,\symRegRange\right)}}}\right)
\\
&+
\sum_{\symRegVal=1}^{\symRegRange+1}
\symCountVariate^{=}_{\symRegVal}
\log\!\left(
1
-e^{-\frac{e^{\symPhi_\symSetASuffix}
+e^{\symPhi_\symSetXSuffix}
}{2^{\min\left(\symRegVal,\symRegRange\right)}}}
-
e^{-\frac{e^{\symPhi_\symSetBSuffix}
+e^{\symPhi_\symSetXSuffix}
}{2^{\min\left(\symRegVal,\symRegRange\right)}}}
+
e^{-\frac{e^{\symPhi_\symSetASuffix}
+e^{\symPhi_\symSetBSuffix}
+e^{\symPhi_\symSetXSuffix}
}{2^{\min\left(\symRegVal,\symRegRange\right)}}}
\right).
\end{aligned}
\end{multline}

Quasi-Newton methods are commonly used for finding the maximum. They require the calculation of the first derivatives which are given by
\begin{align}
\frac{\partial\symFunc}{\partial\symPhi_\symSetASuffix}
=
&
-
\sum_{\symRegVal=0}^{\symRegRange}
\left(
  \symCountVariate^{<}_{1\symRegVal}+
  \symCountVariate^{=}_{\symRegVal}+
  \symCountVariate^{>}_{1\symRegVal}
\right)
\frac{e^{\symPhi_\symSetASuffix}}{2^{\symRegVal}}
+
\sum_{\symRegVal=1}^{\symRegRange}
\symCountVariate^{<}_{1\symRegVal}
\symHelperV\!\left(
\frac{e^{\symPhi_\symSetASuffix}}
{2^{\symRegVal}}
,
\frac{e^{\symPhi_\symSetXSuffix}}
{2^{\symRegVal}}
\right)
+
\sum_{\symRegVal=1}^{\symRegRange+1}
\symCountVariate^{>}_{1\symRegVal}
\symHelperV\!\left(
\frac{e^{\symPhi_\symSetASuffix}}
{2^{\min\left(\symRegVal,\symRegRange\right)}}
,
0
\right)
\nonumber\\
&
+
\sum_{\symRegVal=1}^{\symRegRange+1}
\symCountVariate^{=}_{\symRegVal}
\symHelperW\!
\left(
\frac{e^{\symPhi_\symSetASuffix}}
{2^{\min\left(\symRegVal,\symRegRange\right)}}
,
\frac{e^{\symPhi_\symSetBSuffix}}
{2^{\min\left(\symRegVal,\symRegRange\right)}}
,
\frac{e^{\symPhi_\symSetXSuffix}}
{2^{\min\left(\symRegVal,\symRegRange\right)}}
\right)
\end{align}

\begin{align}
\frac{\partial\symFunc}{\partial\symPhi_\symSetBSuffix}
=
&
-
\sum_{\symRegVal=0}^{\symRegRange}
\left(
  \symCountVariate^{<}_{2\symRegVal}+
  \symCountVariate^{=}_{\symRegVal}+
  \symCountVariate^{>}_{2\symRegVal}
\right)
\frac{e^{\symPhi_\symSetBSuffix}}{2^{\symRegVal}}
+
\sum_{\symRegVal=1}^{\symRegRange}
\symCountVariate^{<}_{2\symRegVal}
\symHelperV\!\left(
\frac{e^{\symPhi_\symSetBSuffix}}
{2^{\symRegVal}}
,
\frac{e^{\symPhi_\symSetXSuffix}}
{2^{\symRegVal}}
\right)
+
\sum_{\symRegVal=1}^{\symRegRange+1}
\symCountVariate^{>}_{2\symRegVal}
\symHelperV\!\left(
\frac{e^{\symPhi_\symSetBSuffix}}
{2^{\min\left(\symRegVal,\symRegRange\right)}}
,
0
\right)
\nonumber\\
&
+
\sum_{\symRegVal=1}^{\symRegRange+1}
\symCountVariate^{=}_{\symRegVal}
\symHelperW\!
\left(
\frac{e^{\symPhi_\symSetBSuffix}}
{2^{\min\left(\symRegVal,\symRegRange\right)}}
,
\frac{e^{\symPhi_\symSetASuffix}}
{2^{\min\left(\symRegVal,\symRegRange\right)}}
,
\frac{e^{\symPhi_\symSetXSuffix}}
{2^{\min\left(\symRegVal,\symRegRange\right)}}
\right)
\end{align}

\begin{align}
\frac{\partial\symFunc}{\partial\symPhi_\symSetXSuffix}
=
&
-
\sum_{\symRegVal=0}^{\symRegRange}
\left(
  \symCountVariate^{<}_{1\symRegVal}+
  \symCountVariate^{=}_{\symRegVal}+
  \symCountVariate^{<}_{2\symRegVal}
\right)
\frac{e^{\symPhi_\symSetXSuffix}}{2^{\symRegVal}}
+
\sum_{\symRegVal=1}^{\symRegRange}
\symCountVariate^{<}_{1\symRegVal}
\symHelperV\!\left(
\frac{e^{\symPhi_\symSetXSuffix}}
{2^{\symRegVal}}
,
\frac{e^{\symPhi_\symSetASuffix}}
{2^{\symRegVal}}
\right)
+
\sum_{\symRegVal=1}^{\symRegRange}
\symCountVariate^{<}_{2\symRegVal}
\symHelperV\!\left(
\frac{e^{\symPhi_\symSetXSuffix}}
{2^{\symRegVal}}
,
\frac{e^{\symPhi_\symSetBSuffix}}
{2^{\symRegVal}}
\right)
\nonumber\\
&
+
\sum_{\symRegVal=1}^{\symRegRange+1}
\symCountVariate^{=}_{\symRegVal}
\symHelperU\!
\left(
\frac{e^{\symPhi_\symSetASuffix}}
{2^{\min\left(\symRegVal,\symRegRange\right)}}
,
\frac{e^{\symPhi_\symSetBSuffix}}
{2^{\min\left(\symRegVal,\symRegRange\right)}}
,
\frac{e^{\symPhi_\symSetXSuffix}}
{2^{\min\left(\symRegVal,\symRegRange\right)}}
\right)
\end{align}

\begin{equation}
\symHelperV\!\left(\symX_1,\symX_2\right)
:=
\frac{\symX_1}{
e^{\symX_1+\symX_2}-1
}
=
\frac{\symX_1}{\symX_1 + \symX_2}
\frac{
\symX_1 + \symX_2
}
{
e^{\symX_1+\symX_2}-1
}
\end{equation}
\begin{equation}
\symHelperW\!\left(\symX_1,\symX_2,\symX_3\right)
:=
\frac{\symX_1\left(e^{\symX_2}-1\right)
}{
e^{\symX_1+\symX_2+\symX_3}-e^{\symX_1}-e^{\symX_2}+1
}
\end{equation}
\begin{equation}
\symHelperU\!\left(\symX_1,\symX_2,\symX_3\right)
:=
\frac{\left(e^{\symX_1}+e^{\symX_2}-1\right)\symX_3
}{
e^{\symX_1+\symX_2+\symX_3}-e^{\symX_1}-e^{\symX_2}+1
}
\end{equation}

>>>>

\documentclass[a4paper]{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{svg}
\usepackage{algpseudocode,algorithm,algorithmicx}
\usepackage{float}
\usepackage{url}
\usepackage{booktabs}
\usepackage{array}
\usepackage{paralist}
\usepackage{verbatim}
\usepackage{subfig}
\usepackage{hyperref}
\usepackage{siunitx}
\usepackage{csvsimple}
\usepackage{multirow}
\usepackage[capitalise,nameinlink,noabbrev]{cleveref}

% SYMBOLS
\newcommand*{\symDefine}[2]{\newcommand{{#1}}{{#2}}}
\symDefine{\symPrecision}{p}
\symDefine{\symRegRange}{q}
\symDefine{\symRegVal}{k}
\symDefine{\symRegValVariate}{K}
\symDefine{\symNumReg}{m}
\symDefine{\symDataItem}{D}
\symDefine{\symBitRepA}{a}
\symDefine{\symBitRepB}{b}
\symDefine{\symIndexI}{i}
\symDefine{\symIndexK}{k}
\symDefine{\symIndexL}{l}
\symDefine{\symIndexJ}{j}
\symDefine{\symCount}{c}
\symDefine{\symCountVariate}{C}
\symDefine{\symAlpha}{\alpha}
\symDefine{\symBeta}{\beta}
\symDefine{\symA}{a}
\symDefine{\symB}{b}
\symDefine{\symKappa}{\kappa}
\symDefine{\symX}{x}
\symDefine{\symY}{y}
\symDefine{\symZ}{z}
\symDefine{\symPhi}{\varphi}
\symDefine{\symXEstimate}{\hat{\symX}}
\symDefine{\symXNormalized}{\symX'}
\symDefine{\symN}{n}
\symDefine{\symS}{s}
\symDefine{\symInteger}{z}
\symDefine{\symIntegerVariate}{Z}
\symDefine{\symCardinality}{n}
\symDefine{\symCardinalityEstimate}{\hat{\symCardinality}}
\symDefine{\symCardinalityRawEstimate}{\symCardinalityEstimate_\text{raw}}
\symDefine{\symError}{\varepsilon}
\symDefine{\symStopDelta}{\delta}
\symDefine{\symStopEpsilon}{\varepsilon}
\symDefine{\symBigO}{\mathcal{O}}
\symDefine{\symExpectation}{\mathbb{E}}
\symDefine{\symProbability}{P}
\symDefine{\symProbabilityMass}{\rho}
\symDefine{\symRegProbability}{\gamma}
\symDefine{\symLikelihood}{\mathcal{L}}
\symDefine{\symPoissonRate}{\lambda}
\symDefine{\symPoissonRateEstimate}{\hat{\symPoissonRate}}
\symDefine{\symFunc}{f}
\symDefine{\symFuncPrime}{g}
\symDefine{\symHelper}{h}
\symDefine{\symHelperApprox}{\tilde{\symHelper}}
\symDefine{\symHelperW}{w}
\symDefine{\symHelperV}{v}
\symDefine{\symHelperU}{u}
\symDefine{\symSetA}{A}
\symDefine{\symSetB}{B}
\symDefine{\symSetS}{S}
\symDefine{\symSetX}{X}
\symDefine{\symSetASuffix}{a}
\symDefine{\symSetBSuffix}{b}
\symDefine{\symSetXSuffix}{x}
\symDefine{\symCountMatrix}{\mathbf{c}}
\symDefine{\symPowerSeriesFunc}{\xi}
\symDefine{\symSmallCorrectionFunc}{\sigma}
\symDefine{\symLargeCorrectionFunc}{\tau}
\symDefine{\symEpsPowerSeriesFunc}{\varepsilon_\symPowerSeriesFunc}

\newcommand{\numformat}[1]{{\num[scientific-notation = true,round-mode = places,round-precision = 3, output-exponent-marker = \ensuremath{\mathrm{e}}]{#1}}}

\newcommand{\numformattwo}[1]{{\num[scientific-notation = fixed,round-mode = places,round-precision = 3]{#1}}}

@@

The transformed log-likelihood function can be written as
%\begin{multline}
%\symFunc(\symPhi_\symSetASuffix,\symPhi_\symSetBSuffix,\symPhi_\symSetXSuffix):=
%\log \symLikelihood(
%\symNumReg e^{\symPhi_\symSetASuffix},
%\symNumReg e^{\symPhi_\symSetBSuffix},
%\symNumReg e^{\symPhi_\symSetXSuffix}
%\vert
%\boldsymbol{\symRegValVariate}_1,
%\boldsymbol{\symRegValVariate}_2
%)=
%\\
%\begin{aligned}
%&-
%\sum_{\symRegVal=0}^{\symRegRange}
%\left(
%  \symCountVariate^{<}_{1\symRegVal}+
%  \symCountVariate^{=}_{\symRegVal}+
%  \symCountVariate^{>}_{1\symRegVal}
%\right)
%\frac{e^{\symPhi_\symSetASuffix}}{2^{\symRegVal}}
%+
%\left(
%  \symCountVariate^{<}_{2\symRegVal}+
%  \symCountVariate^{=}_{\symRegVal}+
%  \symCountVariate^{>}_{2\symRegVal}
%\right)
%\frac{e^{\symPhi_\symSetBSuffix}}{2^{\symRegVal}}
%+
%\left(
%  \symCountVariate^{<}_{1\symRegVal}+
%  \symCountVariate^{=}_{\symRegVal}+
%  \symCountVariate^{<}_{2\symRegVal}
%\right)
%\frac{e^{\symPhi_\symSetXSuffix}}{2^{\symRegVal}}
%\\
%&+
%\sum_{\symRegVal=1}^{\symRegRange}
%\symCountVariate^{<}_{1\symRegVal}
%\log\!\left(1-e^{-\frac{e^{\symPhi_\symSetASuffix}
%+e^{\symPhi_\symSetXSuffix}
%}{2^{\symRegVal}}}\right)
%+
%\symCountVariate^{<}_{2\symRegVal}
%\log\!\left(1-e^{-\frac{e^{\symPhi_\symSetBSuffix}
%+
%e^{\symPhi_\symSetXSuffix}
%}{2^{\symRegVal}}}\right)
%\\
%&+
%\sum_{\symRegVal=1}^{\symRegRange+1}
%\symCountVariate^{>}_{1\symRegVal}
%\log\!\left(1-e^{-\frac{e^{\symPhi_\symSetASuffix}
%}{2^{\min\left(\symRegVal,\symRegRange\right)}}}\right)
%+
%\symCountVariate^{>}_{2\symRegVal}
%\log\!\left(1-e^{-\frac{e^{\symPhi_\symSetBSuffix}
%}{2^{\min\left(\symRegVal,\symRegRange\right)}}}\right)
%\\
%&+
%\sum_{\symRegVal=1}^{\symRegRange+1}
%\symCountVariate^{=}_{\symRegVal}
%\log\!\left(
%1
%-e^{-\frac{e^{\symPhi_\symSetASuffix}
%+e^{\symPhi_\symSetXSuffix}
%}{2^{\min\left(\symRegVal,\symRegRange\right)}}}
%-
%e^{-\frac{e^{\symPhi_\symSetBSuffix}
%+e^{\symPhi_\symSetXSuffix}
%}{2^{\min\left(\symRegVal,\symRegRange\right)}}}
%+
%e^{-\frac{e^{\symPhi_\symSetASuffix}
%+e^{\symPhi_\symSetBSuffix}
%+e^{\symPhi_\symSetXSuffix}
%}{2^{\min\left(\symRegVal,\symRegRange\right)}}}
%\right).
%\end{aligned}
%\end{multline}

\begin{multline}
\symFunc(\symPhi_\symSetASuffix,\symPhi_\symSetBSuffix,\symPhi_\symSetXSuffix):=
\log \symLikelihood(
\symNumReg e^{\symPhi_\symSetASuffix},
\symNumReg e^{\symPhi_\symSetBSuffix},
\symNumReg e^{\symPhi_\symSetXSuffix}
\vert
\boldsymbol{\symRegValVariate}_1,
\boldsymbol{\symRegValVariate}_2
)=
\\
\begin{aligned}
&+
\sum_{\symRegVal=1}^{\symRegRange}
\symCountVariate^{<}_{1\symRegVal}
\log\!\left(
\symZ_{\symSetXSuffix\symRegVal}
+
\symY_{\symSetXSuffix\symRegVal}
\symZ_{\symSetASuffix\symRegVal}
\right)
+
\symCountVariate^{<}_{2\symRegVal}
\log\!\left(
\symZ_{\symSetXSuffix\symRegVal}
+
\symY_{\symSetXSuffix\symRegVal}
\symZ_{\symSetBSuffix\symRegVal}
\right)
\\
&+
\sum_{\symRegVal=1}^{\symRegRange}
\symCountVariate^{>}_{1\symRegVal}
\log\!\left(
\symZ_{\symSetASuffix\symRegVal}
\right)
+
\symCountVariate^{>}_{2\symRegVal}
\log\!\left(
\symZ_{\symSetBSuffix\symRegVal}
\right)
+
\symCountVariate^{=}_{\symRegVal}
\log\!\left(
\symZ_{\symSetXSuffix\symRegVal}
+
\symY_{\symSetXSuffix\symRegVal}
\symZ_{\symSetASuffix\symRegVal}
\symZ_{\symSetBSuffix\symRegVal}
\right)
\\
&+
\symCountVariate^{>}_{1,\symRegRange+1}
\log\!\left(
\symZ_{\symSetASuffix\symRegRange}
\right)
+
\symCountVariate^{>}_{2,\symRegRange+1}
\log\!\left(
\symZ_{\symSetBSuffix\symRegRange}
\right)
+
\symCountVariate^{=}_{\symRegRange+1}
\log\!\left(
\symZ_{\symSetXSuffix\symRegRange}
+
\symY_{\symSetXSuffix\symRegRange}
\symZ_{\symSetASuffix\symRegRange}
\symZ_{\symSetBSuffix\symRegRange}
\right)
\\
&-
\sum_{\symRegVal=0}^{\symRegRange}
\left(
  \symCountVariate^{<}_{1\symRegVal}+
  \symCountVariate^{=}_{\symRegVal}+
  \symCountVariate^{>}_{1\symRegVal}
\right)
\symX_{\symSetASuffix\symRegVal}
+
\left(
  \symCountVariate^{<}_{2\symRegVal}+
  \symCountVariate^{=}_{\symRegVal}+
  \symCountVariate^{>}_{2\symRegVal}
\right)
\symX_{\symSetBSuffix\symRegVal}
+
\left(
  \symCountVariate^{<}_{1\symRegVal}+
  \symCountVariate^{=}_{\symRegVal}+
  \symCountVariate^{<}_{2\symRegVal}
\right)
\symX_{\symSetXSuffix\symRegVal}
.
\end{aligned}
\end{multline}
Here we introduced the following expressions for simplification:
\begin{equation}
\begin{aligned}
\symX_{\symSetASuffix\symRegVal} &:=\frac{e^{\symPhi_\symSetASuffix}}{2^{\symRegVal}},
&
\symY_{\symSetASuffix\symRegVal} &:= e^{-\symX_{\symSetASuffix\symRegVal}},
&
\symZ_{\symSetASuffix\symRegVal} &:= 1 - \symY_{\symSetASuffix\symRegVal},
\\
\symX_{\symSetBSuffix\symRegVal} &:=\frac{e^{\symPhi_\symSetBSuffix}}{2^{\symRegVal}},
&
\symY_{\symSetBSuffix\symRegVal} &:= e^{-\symX_{\symSetBSuffix\symRegVal}},
&
\symZ_{\symSetBSuffix\symRegVal} &:= 1 - \symY_{\symSetBSuffix\symRegVal},
\\
\symX_{\symSetXSuffix\symRegVal} &:=\frac{e^{\symPhi_\symSetXSuffix}}{2^{\symRegVal}},
&
\symY_{\symSetXSuffix\symRegVal} &:= e^{-\symX_{\symSetXSuffix\symRegVal}},
&
\symZ_{\symSetXSuffix\symRegVal} &:= 1 - \symY_{\symSetXSuffix\symRegVal}.
\end{aligned}
\end{equation}

Quasi-Newton methods are commonly used for finding the maximum. They require the calculation of the first derivatives which are given by

\begin{multline}
\frac{\partial\symFunc}{\partial\symPhi_\symSetASuffix}
=
\sum_{\symRegVal=1}^{\symRegRange}
\symCountVariate^{<}_{1\symRegVal}
\frac{
\symY_{\symSetXSuffix\symRegVal}
\symX_{\symSetASuffix\symRegVal}
\symY_{\symSetASuffix\symRegVal}
}{
\symZ_{\symSetXSuffix\symRegVal}
+
\symY_{\symSetXSuffix\symRegVal}
\symZ_{\symSetASuffix\symRegVal}
}
+
\symCountVariate^{=}_{\symRegVal}
\frac{
\symY_{\symSetXSuffix\symRegVal}
\symX_{\symSetASuffix\symRegVal}
\symY_{\symSetASuffix\symRegVal}
\symZ_{\symSetBSuffix\symRegVal}
}
{
\symZ_{\symSetXSuffix\symRegVal}
+
\symY_{\symSetXSuffix\symRegVal}
\symZ_{\symSetASuffix\symRegVal}
\symZ_{\symSetBSuffix\symRegVal}
}
+
\symCountVariate^{>}_{1\symRegVal}
\frac{
\symX_{\symSetASuffix\symRegVal}
\symY_{\symSetASuffix\symRegVal}
}{
\symZ_{\symSetASuffix\symRegVal}
}
\\
+
\symCountVariate^{=}_{\symRegRange+1}
\frac{
\symY_{\symSetXSuffix\symRegRange}
\symX_{\symSetASuffix\symRegRange}
\symY_{\symSetASuffix\symRegRange}
\symZ_{\symSetBSuffix\symRegRange}
}
{
\symZ_{\symSetXSuffix\symRegRange}
+
\symY_{\symSetXSuffix\symRegRange}
\symZ_{\symSetASuffix\symRegRange}
\symZ_{\symSetBSuffix\symRegRange}
}
+
\symCountVariate^{>}_{1,\symRegRange+1}
\frac{
\symX_{\symSetASuffix\symRegRange}
\symY_{\symSetASuffix\symRegRange}
}{
\symZ_{\symSetASuffix\symRegRange}
}
-
\sum_{\symRegVal=0}^{\symRegRange}
\left(
  \symCountVariate^{<}_{1\symRegVal}+
  \symCountVariate^{=}_{\symRegVal}+
  \symCountVariate^{>}_{1\symRegVal}
\right)
\symX_{\symSetASuffix\symRegVal}
\end{multline}

\begin{multline}
\frac{\partial\symFunc}{\partial\symPhi_\symSetBSuffix}
=
\sum_{\symRegVal=1}^{\symRegRange}
\symCountVariate^{<}_{1\symRegVal}
\frac{
\symY_{\symSetXSuffix\symRegVal}
\symX_{\symSetBSuffix\symRegVal}
\symY_{\symSetBSuffix\symRegVal}
}{
\symZ_{\symSetXSuffix\symRegVal}
+
\symY_{\symSetXSuffix\symRegVal}
\symZ_{\symSetBSuffix\symRegVal}
}
+
\symCountVariate^{=}_{\symRegVal}
\frac{
\symY_{\symSetXSuffix\symRegVal}
\symX_{\symSetBSuffix\symRegVal}
\symY_{\symSetBSuffix\symRegVal}
\symZ_{\symSetASuffix\symRegVal}
}
{
\symZ_{\symSetXSuffix\symRegVal}
+
\symY_{\symSetXSuffix\symRegVal}
\symZ_{\symSetBSuffix\symRegVal}
\symZ_{\symSetASuffix\symRegVal}
}
+
\symCountVariate^{>}_{1\symRegVal}
\frac{
\symX_{\symSetBSuffix\symRegVal}
\symY_{\symSetBSuffix\symRegVal}
}{
\symZ_{\symSetBSuffix\symRegVal}
}
\\
+
\symCountVariate^{=}_{\symRegRange+1}
\frac{
\symY_{\symSetXSuffix\symRegRange}
\symX_{\symSetBSuffix\symRegRange}
\symY_{\symSetBSuffix\symRegRange}
\symZ_{\symSetASuffix\symRegRange}
}
{
\symZ_{\symSetXSuffix\symRegRange}
+
\symY_{\symSetXSuffix\symRegRange}
\symZ_{\symSetBSuffix\symRegRange}
\symZ_{\symSetASuffix\symRegRange}
}
+
\symCountVariate^{>}_{1,\symRegRange+1}
\frac{
\symX_{\symSetBSuffix\symRegRange}
\symY_{\symSetBSuffix\symRegRange}
}{
\symZ_{\symSetBSuffix\symRegRange}
}
-
\sum_{\symRegVal=0}^{\symRegRange}
\left(
  \symCountVariate^{<}_{1\symRegVal}+
  \symCountVariate^{=}_{\symRegVal}+
  \symCountVariate^{>}_{1\symRegVal}
\right)
\symX_{\symSetBSuffix\symRegVal}
\end{multline}

\begin{multline}
\frac{\partial\symFunc}{\partial\symPhi_\symSetXSuffix}
=
\sum_{\symRegVal=1}^{\symRegRange}
\symCountVariate^{<}_{1\symRegVal}
\frac{
\symY_{\symSetXSuffix\symRegVal}
\symY_{\symSetASuffix\symRegVal}
\symX_{\symSetXSuffix\symRegVal}
}{
\symZ_{\symSetXSuffix\symRegVal}
+
\symY_{\symSetXSuffix\symRegVal}
\symZ_{\symSetASuffix\symRegVal}
}
+
\symCountVariate^{=}_{\symRegVal}
\frac{
\symX_{\symSetXSuffix\symRegVal}
\symY_{\symSetXSuffix\symRegVal}
\left(
\symZ_{\symSetASuffix\symRegVal}
\symY_{\symSetBSuffix\symRegVal}
+
\symY_{\symSetASuffix\symRegVal}
\symZ_{\symSetBSuffix\symRegVal}
+
\symY_{\symSetASuffix\symRegVal}
\symY_{\symSetBSuffix\symRegVal}
\right)
}
{
\symZ_{\symSetXSuffix\symRegVal}
+
\symY_{\symSetXSuffix\symRegVal}
\symZ_{\symSetBSuffix\symRegVal}
\symZ_{\symSetASuffix\symRegVal}
}
+
\symCountVariate^{<}_{2\symRegVal}
\frac{
\symY_{\symSetXSuffix\symRegVal}
\symY_{\symSetBSuffix\symRegVal}
\symX_{\symSetXSuffix\symRegVal}
}{
\symZ_{\symSetXSuffix\symRegVal}
+
\symY_{\symSetXSuffix\symRegVal}
\symZ_{\symSetBSuffix\symRegVal}
}
\\
+
\symCountVariate^{=}_{\symRegRange+1}
\frac{
\symX_{\symSetXSuffix\symRegRange}
\symY_{\symSetXSuffix\symRegRange}
\left(
\symZ_{\symSetASuffix\symRegRange}
\symY_{\symSetBSuffix\symRegRange}
+
\symY_{\symSetASuffix\symRegRange}
\symZ_{\symSetBSuffix\symRegRange}
+
\symY_{\symSetASuffix\symRegRange}
\symY_{\symSetBSuffix\symRegRange}
\right)
}
{
\symZ_{\symSetXSuffix\symRegRange}
+
\symY_{\symSetXSuffix\symRegRange}
\symZ_{\symSetBSuffix\symRegRange}
\symZ_{\symSetASuffix\symRegRange}
}
-
\sum_{\symRegVal=0}^{\symRegRange}
\left(
  \symCountVariate^{<}_{1\symRegVal}+
  \symCountVariate^{=}_{\symRegVal}+
  \symCountVariate^{<}_{2\symRegVal}
\right)
\symX_{\symSetXSuffix\symRegVal}
\end{multline}
